Parsing Postfix log files

### 2009-03-26 doc/logparser-results.tex
Track the number of lines per program too.
Changes from proofreading, and a couple of other checks.

XXX USE THE DATA COLLECTED ABOUT NUMBER OF RULES TRIED\@.  MAYBE INCLUDE IN
ORDERINGS SECTION\@?  MIX WITH RECOGNISING VS RECOGNISING + ACTION, AND
KEEP COUNTERS FOR EACH PROGRAM TOO --- MAYBE SOMETHING USEFUL WILL COME
FROM IT\@.

@@

At the time of writing there are \numberOFrules{} different rules: the top
10\% match the vast majority of log lines, with the remaining log lines
split across the other 90\% of the rules (as shown in \graphref{rule hits
graph}).  Assuming that the distribution of log lines is reasonably
consistent over time, parser efficiency should benefit from trying more
frequently matching rules before those that match less frequently.  To
test this hypothesis three full test runs were performed with different
rule orderings:

@@

    \item [Optimal]  The most optimal order, according to the hypothesis:
        rules that match most often will be tried first.

@@

    \item [Reverse] Hypothetically the worst order: the most frequently
        matching rules will be tried last.

@@

It is clear from \graphref{Percentage parsing time increase of maximum
ruleset over minimum ruleset} that the increased number of rules noticeably
increases parsing time with reverse ordering, and to a lesser extent with
shuffled ordering.  The optimal ordering (where the most frequently
matching rules are tried first) shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  Log
files 22 and 62--68 show much smaller increases for shuffled and reversed
ordering than other log files do, because the majority of log lines in
those log files are parsed by Postfix components with a small number of
parsing rules, so removing unnecessary rules has little effect on the total
number of rules used.  These results show that the architecture scales very
well as the number of rules increases, and that optimally sorting the rules
is an important optimisation contributing to this scalability.

@@

    \item Try each rule until a matching rule is found; if no match is
        issue a warning and move to the next log line.

@@

    \item There must be a rule to match each different log line produced by
        each program; if a log line is not successfully matched the parser
        will issue a warning.  Rules should be as specific and tightly
        bound as possible to ensure accurate parsing:\footnote{A rule that
        matches zero or more of any character will successfully parse every
        log line, but not in a meaningful way.} most log lines contain
        fixed strings and have a rigid pattern, so this is not a problem.

@@

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILESall{} contiguous log files from Postfix 2.2 through to
Postfix 2.5), and new rules are easy to add.  Maintaining full coverage is
hard because other servers have different restrictions with custom
messages, \acronym{DNSBL} messages change over time, major releases of
Postfix change warning messages (usually adding more information), etc.,\
so over time the log lines drift and change.  \Graphref{rule hits graph}
shows the number of hits for each rule over all \numberOFlogFILES{} log
files.  The number of hits per rule is quite unevenly spread, resembling a
Power Law distribution --- it is obvious that a small number of rules match
the vast majority of the log lines, and more than half the rules match
fewer than 100 times.

>>>>

XXX USE THE DATA COLLECTED ABOUT NUMBER OF RULES TRIED\@.  MAYBE INCLUDE IN
ORDERINGS SECTION\@?  MIX WITH RECOGNISING VS RECOGNISING + ACTION, AND
KEEP COUNTERS FOR EACH PROGRAM TOO --- MAYBE SOMETHING USEFUL WILL COME
FROM IT\@.  COMPARE NUMBER OF RULES TRIED FOR DIFFERENT RULE ORDERINGS\@.

@@

At the time of writing there are \numberOFrules{} different rules: the top
10\% recognise the vast majority of log lines, with the remaining log lines
split across the other 90\% of the rules (as shown in \graphref{rule hits
graph}).  Assuming that the distribution of log lines is reasonably
consistent over time, parser efficiency should benefit from trying more
frequently recognising rules before those that recognise less frequently.
To test this hypothesis three full test runs were performed with different
rule orderings:

@@

    \item [Optimal]  The most optimal order, according to the hypothesis:
        rules that recognise log lines most often will be tried first.

@@

    \item [Reverse] Hypothetically the worst order: the rules that
        recognise log lines most frequently will be tried last.

@@

It is clear from \graphref{Percentage parsing time increase of maximum
ruleset over minimum ruleset} that the increased number of rules noticeably
increases parsing time with reverse ordering, and to a lesser extent with
shuffled ordering.  The optimal ordering (where the rules that recognise
log lines most frequently are tried first) shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  Log
files 22 and 62--68 show much smaller increases for shuffled and reversed
ordering than other log files do, because the majority of log lines in
those log files are parsed by Postfix components with a small number of
parsing rules, so removing unnecessary rules has little effect on the total
number of rules used.  These results show that the architecture scales very
well as the number of rules increases, and that optimally sorting the rules
is an important optimisation contributing to this scalability.

@@

    \item Try each rule until a recognising rule is found; if the log line
        is not recognised, issue a warning and move to the next log line.

@@

    \item There must be a rule to recognise each different log line
        produced by each program; if a log line is not successfully
        recognised the parser will issue a warning.  Rules should be as
        specific and tightly bound as possible to ensure accurate
        parsing:\footnote{A rule with a regex that matches zero or more of
        any character will successfully recognise every log line, but not
        in a meaningful way.} most log lines contain fixed strings and have
        a rigid pattern, so this is not a problem.

@@

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILESall{} contiguous log files from Postfix 2.2 through to
Postfix 2.5), and new rules are easy to add.  Maintaining full coverage is
hard because other servers have different restrictions with custom
messages, \acronym{DNSBL} messages change over time, major releases of
Postfix change warning messages (usually adding more information), etc.,\
so over time the log lines drift and change.  \Graphref{rule hits graph}
shows the number of hits for each rule over all \numberOFlogFILES{} log
files.  The number of hits per rule is quite unevenly spread, resembling a
Power Law distribution --- it is obvious that a small number of rules
recognise the vast majority of the log lines, and more than half the rules
recognise fewer than 100 log lines.
